---
title: "Homework 12"
author: "Stefan Konigorski"
date: "January 25, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Linear mixed model for hierarchical data

Load the HRS dataset and do the following preparations:

```{r}
library(readr)
dat_link <- url("https://www.dropbox.com/s/z281ijk81ohyg9v/hrs.RData?dl=1")
load(dat_link)

# Preparations
# rename variable at first timepoint
names(hrs)[5] <- "R1IMRC"
# define variable at second timepoint - Integrate the variables R2AIMRC10 and R2HIMRC20, which both capture different subgroups of the sample at the second timepoint.
table(is.na(hrs$R2AIMRC10)) # -> only ~7000 of ~36000 observations
table(is.na(hrs$R2HIMRC20)) # -> only ~10000 of ~36000 observations
table(!is.na(hrs$R2HIMRC20) & !is.na(hrs$R2AIMRC10)) # -> no person has a value at both variables, i.e. the variables complement each other
hrs$R2IMRC <- rowSums( cbind (hrs$R2AIMRC10, hrs$R2HIMRC20), na.rm=TRUE)
hrs$R2IMRC[!(!is.na(hrs$R2HIMRC20) | !is.na(hrs$R2AIMRC10))] <- NA
table(is.na(hrs$R2IMRC))
# Rescale R1IMRC and R2IMRC onto the same scale as all other variables (measures 3-12 were on a scale of 1-10, measures 1-2 on the scale 1-20)
hrs$R1IMRC <- hrs$R1IMRC/2
hrs$R2IMRC <- hrs$R2IMRC/2

# Extract variables for first investigation
data_IMRC <- data.frame(HHID = hrs$HHID, PN = hrs$PN, HHID_PN = paste(hrs$HHID, hrs$PN, sep = "_"),
                        T1 = hrs$R1IMRC, T2 = hrs$R2IMRC, T3 = hrs$R3IMRC, T4 = hrs$R4IMRC, T5 = hrs$R5IMRC, T6 = hrs$R6IMRC,
                        T7 = hrs$R7IMRC, T8 = hrs$R8IMRC, T9 = hrs$R9IMRC, T10 = hrs$R10IMRC, T11 = hrs$R11IMRC, T12 = hrs$R12IMRC)
```

Answer the following questions: 


a) How many observations are there in the dataset?
## There are 36167 observations in the dataset.

b) How many households are there in the dataset (observations are clustered in households)?
```{r}
length(unique(data_IMRC$HHID))
```
## There are 22996 households in the dataset.

c) Now, compute a linear regression model to predict T2 by T1, i.e. the cognitive performance at the second timepoint by the cognitive performance at the first timepoint. Also, compute two linear mixed models, one with random intercept, one with random intercept+slope to predict T2 by T1. Compare the results, i.e. the estimated effect of T1 on T2. Are they similar, or are they very different?

## computing linear regression model to predict T2 by T1

```{r}
fit1 <- lm(T2 ~ T1, data = data_IMRC)
summary(fit1)
pred.fit1 <- predict(fit1)
head(pred.fit1)
```

## computing linear mixed models, with random intercept

```{r}
library(nlme)
fit2 <- lme(T2 ~ T1, data = data_IMRC,na.action = na.exclude, random = ~1| HHID)
summary(fit2)
pred.fit2 <- predict(fit2)
head(pred.fit2)
```
## Computing linear mixed models, with random intercept+slope

```{r}
library(nlme)
fit3 <- lme(T2 ~ T1, data = data_IMRC,na.action = na.exclude, random = ~ T1| HHID)
summary(fit3)
pred.fit3 <- predict(fit3)
head(pred.fit3)
```
## Results:
The first model is very inaccurate as the prediction is very far away from the actual values. In the linear mixed models the prediction is better than the first one yet not close to real values. Adding the slope to the model resulted better and the best out of three. But all of these three models are not so different as they predicted very similar values. 

