---
title: "Homework 3"
author: "Stefan Konigorski"
date: "November 2, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Compute frequencies in the Pima diabetes dataset

Load the Pima diabetes dataset:

```{r}
load(file = url("https://www.dropbox.com/s/hxn1c7yw9r8aw8v/Pima_diabetes.RData?dl=1"))
```

Which variables are measured on a nominal level?

Now compute frequency tables, barplots, and mosaic plots of all nominal variables in the dataset.

```{r}
#frequency tables
# absolute frequencies
table(Pima_diabetes$Outcome)
# relative frequencies
n <- nrow(Pima_diabetes)
table(Pima_diabetes$Outcome) / n
```
```{r}
#Barplots
library(ggplot2)
ggplot(Pima_diabetes, aes(Outcome)) + geom_bar()
```
```{r}
#Mosaic Plot
library(ggmosaic)
Pima_diabetes$Outcomex <- ifelse(Pima_diabetes$Outcome==0,"False","True")
ggplot(data = Pima_diabetes) +
  geom_mosaic(aes(x = product(Outcomex), fill=Outcomex)) + 
  labs(title='Result of Diabetes')
```

Next, create a variable which describes whether a woman had more or less than 4 pregnancies. Then, use this variable to create a 2x2 table with diabetes outcome. Do you see an indication of whether the number of pregnancies is associated with diabetes prevalence? Do you think your investigation is a good way to investigate this?

```{r}
Pima_diabetes$pregx <- ifelse(Pima_diabetes$Pregnancies>4, 0, 1)
table(Pima_diabetes$pregx, Pima_diabetes$Outcome)
```

```{r}
print("So the women who had more than 4 pregnencies were more positive for diabetes.")
```



## Exercise 2: Generate a table with descriptive statistics (optional, but recommended)

Use any dataset (a dataset that you have worked with in the past, or that you are currently working with, a dataset that is available on Blackboard, in R or that you have downloaded from the internet), and generate a table with descriptive statistics of the main variables of interest.

## Exercise 3: Plots using ggplot2

Load the NoShow dataset:

```{r}
load(file = url("https://www.dropbox.com/s/4oqg79cn1qfnhsh/NoShowdata.RData?dl=1"))
head(NoShowdata)
```

Use ggplot2 to generate the following plots:

- Create a boxplots of Age (stratified) by neighborhood.


```{r}
boxes=ggplot(data=NoShowdata, mapping=aes(x=Age, y=Neighbourhood))+geom_boxplot()
boxes
```


- Create a histogram of Age.

```{r}
hists=ggplot(data=NoShowdata, mapping=aes(x=Age))+geom_histogram(binwidth = 2, colour="black", fill="white")
hists
```

- Create a histogram of Age, stratified by whether the person showed up - in one panel using the or in multiple panels.

```{r}
histx=ggplot(NoShowdata, aes(x = Age, color = `No-show`)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 60)
histx

```


- Stratify this plot further by gender.

```{r}
histxy=ggplot(NoShowdata, aes(x = Age, fill = `No-show`)) +
  geom_histogram(aes(color= Gender), position = "identity", alpha = 0.1, bins = 60)
histxy
```


What insights can you get from these plots? For which goal would you create these plots?

## From the boxplot, we can see the sample age summary where we can see the range and denses of all the neighbourhoods but as the number of neighbourhoods are so many, the plot is not that practical. 
## From the histogram we can get the idea of different aged people and then which aged people showed up and not where we see that after the age of 60, the number declines.
## After that we get the insight of the gender differences for the different aged people who showed up or not where we see that female people showed up more than male.