---
title: "Homework 5"
author: "Stefan Konigorski"
date: "November 16, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Probability distributions

Explore the functions *rnorm*, *rt*, *runif*, *rbinom* in R that allow you to generate random numbers from the normal, t-, uniform, and binomial distribution. Compute them with different values, and inspect histograms to visualize their distribution.
```{r}
set.seed(404)
dfX <- rnorm(1000,0,1)
var(dfX)

var_y <- var_z <- NULL
for(i in 1:2000){
  dfX_s <- sample(dfX, size = 100, replace = FALSE)
  var_y[i] <- var(dfX_s)*(100-1)/100
  var_z[i] <- var(dfX_s)
}

# Results in terms of mean
mean(var_y)
mean(var_z)

# Visualizing the histogram
par(mfrow = c(1,2))
hist(var_y)
hist(var_z)
```
## Binomial Distribution
```{r}
set.seed(404)
B <- rbinom(1000, 90, 0.9) # with size = 20, prob=0.2
var(B)

emean <- emedian <- NULL # estimated mean and median initiating
for(i in 1:2000){
  B_sam <- sample(B, size = 100, replace = FALSE)
  emean[i] <- mean(B_sam)
  emedian[i] <- median(B_sam)
}

# mean of variance of estimated mean and median
mean((emean - mean(B))^2)
mean((emedian - mean(B))^2)

var(emean)
var(emedian)

# Visualizing histogram
par(mfrow = c(1,2))
hist(emean)
hist(emedian)
```
##t-distribution
```{r}
set.seed(404)
size = 20000
T <- rt(n=size, df=size-1)
var(T)
hist(T, breaks = 'Scott', freq = FALSE)

emean_t <- emedian_t <- NULL
for(i in 1:2000){
  T_sam <- sample(T, size = 100, replace = FALSE)
  emean_t[i] <- mean(T_sam)
  emedian_t[i] <- median(T_sam)
}
#results
mean((emean_t - mean(T))^2)
mean((emedian_t - mean(T))^2)

#visualization
par(mfrow = c(1,2))
hist(emean_t)
hist(emedian_t)

```
## Continuous Uniform Distribution
```{r}
set.seed(404)
# generating random variables
C <- runif(20000, min = -1, max = 1)
hist(C, freq = FALSE, xlab = 'x', density = 90)

emean_c <- emedian_c <- NULL
for(i in 1:1500){
  C_sam <- sample(C, size = 100, replace = FALSE)
  emean_c[i] <- mean(C_sam)
  emedian_c[i] <- median(C_sam)
}
#results
mean((emean_c - mean(C))^2)
mean((emedian_c - mean(C))^2)

#visualization
par(mfrow = c(1,2))
hist(emean_c)
hist(emedian_c)
```

## Exercise 2: Odds ratio

In the KiGGS dataset:

a) Compute the proportion of mothers that had hypertension during pregnancy. Use the variable 'e0155' which has values "Ja" (yes), "Nein" (No) and "Weiß nicht" (don't know).
```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)

dat <- KiGGS03_06

library(dplyr)
with_hyper <- dat %>% filter(e0155 == "Ja") %>% nrow()
without_na <- sum(!is.na(dat$e0155))
proportion <- (with_hyper/without_na)
proportion
```
b) Create a new variable that is 0 or 1 depending on whether the children are small or tall (think of a good way how to do this) based on the variable 'groeB' (body height). 
```{r}

dat$groeB[is.na(dat$groeB)]= 0
m <- mean(dat$groeB)
dat$v <- ifelse(dat$groeB < m, 0, 1)

```
c) Then compute the odds ratio that the mother had hypertension during pregnancy (e0155 == "Ja" (yes), versus e0155 == "Nein" (no)) of tall vs. small children.
```{r}
dat$groeB[is.na(dat$groeB)]= 0
m <- mean(dat$groeB)
dat$height <- ifelse(dat$groeB < m, 0, 1)

table(dat$height == 0 & dat$e0155 == "Ja")

```

## Exercise 3 (optional): Confidence intervals

Look at the hypertension variable from exercise 2. Use the binom::binom.confint and the questionr::odds.ratio functions to compute the estimates of the proportion and odds ratio as well as their confidence intervals. (you need to download and load these packages at first).

## Exercise 4 (optional, advanced): Bootstrap

Adapt the bootstrap implementation in R_5b_estimation_bootstrap.Rmd to compute the bootstrap estimate of the standard error of the variance of a normally-distributed and a t-distributed variable. Are they similar?