---
title: "Homework 6"
author: "Stefan Konigorski"
date: "November 23, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Statistical tests

a) In the KiGGS dataset, perform a 2-sample t-test for paired samples (i.e. dependent samples) for the two variables 'sys1' and 'sys2'. Interpret the results.

```{r, warning=FALSE, message=FALSE}
# loading all the necessary libraries
library(ggplot2)
library(dplyr)

# importing dataset
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)

dat <- KiGGS03_06

# creating subset for sys1 and sys2
x = matrix(dat$sys1)
y = matrix(dat$sys2)

```

## Now we need to replace the NA values before computing t-test


```{r}
x <- x %>% replace(is.na(.), 0)
y <- y %>% replace(is.na(.), 0)
```


```{r}
# Compute t-test

res <- t.test(as.integer(x), as.integer(y), paired = TRUE)
res
```
### *Here the p-value of the test is less than 2.2e-{16},<br /> which is less than the significance level alpha = 0.05.<br /> We can then reject null hypothesis and conclude that<br /> the sys1 is significantly different from the sys2 with a p-value < 2.2e-{16}.<br /> In other words, we do have sufficient evidence to say that the mean scores between the sys1 and sys2 groups variable is statistically significantly different.<br /> In addition, our 95% confidence interval says that we are “95% confident” that the true mean difference between the two groups is between 1.734622 and 2.058122. So we can reject the null hypothesis here.*

b) In the KiGGS dataset, select one metric and one binary variable (or create one) and perform a 2-sample t-test. Then do a Mann-Whitney U-test with the same variables, explore the function to do this, and compare the results to the t-test.

```{r}
# taking one metric variable m and one binary variable n
m = matrix(dat$wKiGGS)
n = matrix(dat$sex)


```
## Now we need to replace the NA values before computing t-test


```{r}
m <- m %>% replace(is.na(.), 0)
n <- n %>% replace(is.na(.), 0)
```
### Creating dataframe with two variable

```{r}
my_data <- data.frame( 
                metri = m,
                binar = n
                )
```


### Now computing 2 sample t-test for these two variable

```{r}
res <- t.test(metri ~ binar, data = my_data, var.equal = TRUE)
res
```
### Now computing mann-whitney u-test for the same variable with the data frame
```{r}
#calculating the Mann Whitney U test
mann <- wilcox.test(metri ~ binar, data = my_data)
mann
```

### *Here we can see two different scores from t-test and mann whitney u-test on the data frame of one metric variable and one binary variable.<br /> According to t-test we can not reject the null hypothesis as the p value is larger than the significance value 0.05.<br /> On the other hand, mann whitney u-test is showing that we can reject the null hypothesis because of its lower P value than the significance value.<br /> Here, The metric variable data is heavily right skewed which may cause higher error for t-test but the dataset is large for it's advantage where the u-test might suffer from the size.*

```{r}
hist(m, prob = TRUE, main = "Right Skewed Data", xlab = "Metric Variable", col = 'white', breaks = 100)
z <- seq(min(m), max(m), length = 50)
f <- dnorm(z, mean = mean(m), sd = sd(m))
lines(z, f, col = "black", lwd = 1)
```

## Exercise 2: Study planning (optional)

Compute the sample size using R or G*Power, for a study that investigates the question if biking to the HPI is associated with concentration in class or not.

Think about how you can formulate this study question in terms of variables, and which statistical test (that we have covered in class 6) would be appropriate for this.

a) Which test could you use?

b) Which sample size is necessary, to find a true effect with 80\% power at $\alpha=0.05$ using this statistical test?
