---
title: "Homework 7"
author: "Stefan Konigorski"
date: "November 30, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
kiggs <- KiGGS03_06
```

## Exercise 1: Multiple imputation

a) In the KiGGS dataset, choose 2 metric variables of your choice.
```{r}
##  Choosing 2 metric variables
x = kiggs$wKiGGS
y = kiggs$wKGSLab
```

b) Compute the correlation between these two variables

### here we will compute correlation between them.
```{r}
cor_res <- cor.test(x,y, method = "pearson", use = "complete.obs")
cor_res
```
### From the p-value we can see that the variables are significantly correlated with positive correlation coefficient of 0.864. We can generate a plot to visualize it.

```{r}
# plot
library(ggstatsplot)
ggscatterstats(data = kiggs, x = wKiGGS, y = wKGSLab)
```


c) Perform a multiple imputation using the mice function with the 'pmm' method and otherwise default settings.
```{r}
dat <- data.frame(kig = kiggs$wKiGGS, 
                  kiglab = kiggs$wKGSLab)

# using vim to visualize the missing value pattern
library(VIM)
aggr(dat, col = c('skyblue','black'), numbers = TRUE, sortVars = TRUE, labels = names(dat), cex.axis = 0.7, gap = 3, ylab = c("Histogram of missing data", "Pattern"))

```


```{r}
library(mice)

# imputation of missing values, generating 5 imputed data sets
impData <- mice(dat, m = 5, maxit = 5, meth = 'pmm', seed = 404)
?mice

```

```{r}
# Summary of the imputation
summary(impData)
```

### So there are no missing values in the 1st variable but there are many missing values in the other one which is now imputed.

```{r}
# Visualization of the distribution of the imputed variable wkgslab
mice::densityplot(impData)
```

d) Compute the correlation again on the multiple imputed datasets.

```{r}
with(impData, cor.test(kig, kiglab, method = "pearson"))
```
### Here we can see that the correlation is not changed much in the imputed datasets. This might happen as one of the variable has no missing values.
