---
title: "Additional Homework 9"
author: "Stefan Konigorski"
date: "December 14, 2022"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Assumptions of linear regression

Load the KiGGS dataset and compute a regression predicting BMI by sex and age groups (age2):

```{r}
# load data
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06

# Regression:
fit1 <- lm(dat$bmiB ~ dat$sex + as.numeric(dat$age2))

# results:
summary(fit1)
```
## Predicting bmi
```{r}
pred_res <- predict(fit1)
head(pred_res)
head(dat$bmiB)
```


In this model, investigate and judge whether the assumptions listed on slide 13 in lecture 9 are satisfied.


# Checking if bmi is continious
```{r}
library(plyr)
is.discrete(dat$bmiB)
```
## SO the bmi is continious

## Checking linearity
```{r}
plot(fit1,1)
```
## The Plot shows that the linearity exists between the regression line and the residual values.

## Checking if all relevant variables are in the model

```{r}
ndat <- select_if(dat, is.numeric)
cor_list = cor(ndat)
cor_list['bmiB']
```
## So, we can see that bmi has no correlation with any metric values. There might be some ordinal values which has correaltion with age and sex variable

## Checking if the observations are independent
```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")
mdata = data.frame(dat$bmiB, as.numeric(dat$sex), as.numeric(dat$age2))
corz = rquery.cormat(mdata)
print(corz$r)
print(corz$p)
```
## Here we can see that the observations are independent.


## Checking the presence of multicollinearity

## From the previous plot we can see that the predictors have no strong correlation between them. So there is no multicollinearity. 

## Checking homoscedasticty
```{r}
plot(fit1,3)
```
## We can see that the residuals are close to equally distrbuted with a horizontal line to be largely equal in variance with a slight increase towards the right of the plot.

## Checking normal distribution of the residuals
```{r}
plot(fit1,2)
```
## The plot shows that the residuals are along with the straight line which suggests that the residuals are not close to normal distribution.


## Exercise 2: Model selection in linear regression (optional)

In the KiGGS dataset, aim to select relevant predictors for sys12 (systolic blood pressure). Use 2 of the model selection approaches described on slide 26, apply them to the KiGGS dataset and compare the results.

## Exercise 3: Linear regression with multiple imputation (optional)

Run the code in the Rmd file R_9b_linear_regression_MI.Rmd, inspect the R code what it is doing, and look at the results. Apply the same to the linear regression model of another variable of your choice.
